Active-contours driven registration method for structure-informed segmentation of diffusion MR images

### 2015-04-10 Publications/2015-NeuroImage/02-abstract.tex
regseg label

% -*- root: 00-main.tex -*-
We propose a simultaneous segmentation and registration method called \emph{regseg}
  that uses active contours without edges extracted from structural images.
The contours evolve through a free-deformation field of B-Spline functions that
  optimally maps the surfaces onto the data in the target space.
\emph{Regseg} is intended, but not limited, to correct susceptibility-derived distortions
  in echo-planar imaging.
We proof the concept behind \emph{regseg} on four digital phantoms warped with random
  deformations, and conclude that the precision of the segmentation is subvoxel.
Finally, we demonstrate \emph{regseg} in the correction task on 16 real diffusion MRI
  datasets from the \acrlong*{hcp}.
The results suggest that \emph{regseg} overperformed an extended competing method.
We compute the geometrical disagreement of the surfaces registered with \emph{regseg} with
  respect to a ground truth, obtaining a 95\% confidence interval of $[1.08 - 1.50]mm$ in images 
  of $1.25mm$ isotropic resolution.

>>>>

% -*- root: 00-main.tex -*-
Current applications of whole-brain tractography on \acrlong*{dmri} data require highly precise
  delineations of anatomical structures, usually projected from anatomical \acrlong*{t1} images
  through registration.
We propose \regseg{}, a simultaneous segmentation and registration method
  that uses active contours without edges extracted from structural images.
The contours evolve through a free-form deformation field supported by B-Spline basis
  optimally mapping the contours onto the data in the target space.
We tested the functionality of \regseg{} on four digital phantoms warped with known and
  randomly generated deformations, to conclude that the accuracy achieved is subvoxel.
We then apply \regseg{} to perform registration/segmentation task on 16 real diffusion MRI
  datasets from the \acrlong*{hcp}, warped with realistic and nonlinear distortions typically
  present in these data.
We compute the misregistration error of the contours estimated by \regseg{} with respect
  to their theoretical location using the ground-truth, obtaining a 95\% CI of 0.56 - 0.66 (mm)
  distance between corresponding mesh vertices, below the 1.25mm isotropic resolution of the images.
We also compared the performance to a widely-used registration tool, to find out
  that \regseg{} overperformed the alternate method in our settings.

### 2015-04-10 Publications/2015-NeuroImage/03-introduction.tex
regseg label

The hypothesis under this research is that the registration and segmentation
  problems in \gls*{dmri} can be solved simultaneously, increasing the geometrical
  accuracy of the process.
We propose a tool called \emph{regseg} to exploit the prior information of shapes
  extracted on \gls*{t1} images to register the anatomical reference
  to \gls*{dmri} space, implicitly segmenting these data.
The approach uses active contours without edges \citep{chan_active_2001} that evolve driving a
  free-form deformation field of B-Spline basis functions.
Optimization is performed using a descent strategy of explicit shape-gradients
  \citep{besson_dream2s_2003,herbulot_segmentation_2006}.
Hence, unlike the most of presented methods, \emph{regseg} does not implement levelsets.
Since the nonlinear distortion is aligned with one of the imaging axis (see 
  \autoref{sec:human_connectome}), \emph{regseg} includes an anisotropic regularizer of
  the displacements field proposed by \cite{nagel_investigation_1986}.
Finally, we evaluate \emph{regseg} using an extension of our instrumentation framework
  \citep{esteban_simulationbased_2014} that simulates known and realistic distortions
  on \gls*{dmri} data.
The evaluation includes a comparison to the \gls*{t2b} correction by integrating in the framework
  an in-house implementation of the method.

>>>>

The hypothesis under this research is that the registration and segmentation
  problems in \gls*{dmri} can be solved simultaneously, increasing the geometrical
  accuracy of the process.
We propose a tool called \regseg{} to exploit the prior information of shapes
  extracted on \gls*{t1} images to register the anatomical reference
  to \gls*{dmri} space, implicitly segmenting these data.
The approach uses active contours without edges \citep{chan_active_2001} that evolve driving a
  free-form deformation field of B-Spline basis functions.
Optimization is performed using a descent strategy of explicit shape-gradients
  \citep{besson_dream2s_2003,herbulot_segmentation_2006}.
Hence, unlike the most of presented methods, \regseg{} does not implement levelsets.
Since the nonlinear distortion is aligned with one of the imaging axis (see 
  \autoref{sec:human_connectome}), \regseg{} includes an anisotropic regularizer of
  the displacements field proposed by \cite{nagel_investigation_1986}.
Finally, we evaluate \regseg{} using an extension of our instrumentation framework
  \citep{esteban_simulationbased_2014} that simulates known and realistic distortions
  on \gls*{dmri} data.
The evaluation includes a comparison to the \gls*{t2b} correction by integrating in the framework
  an in-house implementation of the method.

### 2015-04-10 Publications/2015-NeuroImage/04-methods.tex
updates and standalone figures
add missing comma after formula
regseg label

 	\begin{align}
  P( M \mid U) &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod}
  \mathcal{N} ( \vec{f} \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} ) = \notag\\
  &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod} \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{l}\right|}}\,{e^{\left(-\frac{1}{2}
  \mdist{f'}{l} \right)}},
  \label{eq:pdf}
  \end{align}
%
  using $\mdist{f}{l}$ to denote the squared \emph{Mahalanobis distance} of $\vec{f}$ with respect
  to the descriptors of region $l$ as
  $\mdist{f}{l} = (\vec{f} - \boldsymbol{\mu}_l)^T \, {\boldsymbol{\Sigma}_l}^{-1} \, (\vec{f} - \boldsymbol{\mu}_l)$.
$C$ is the number of channels comprised in the image $M$.
In fact, the projection of $\gammaset_R$ onto $M$ is an implicit segmentation model, for which
  the covariance matrix $\boldsymbol{\Sigma}_l$ of each region is minimized.

@@

  \begin{align}
  P(U) = \underset{\vec{r}}{\prod}\, p(\vec{u}) &=
  \underset{\vec{r}}{\prod}\, p_0(\vec{u}) \, p_1(\vec{u}), \\
  p_0(\vec{u}) &= \mathcal{N}( \vec{u} \mid 0, \mathbf{A}^{-1}), \notag\\
  p_1(\vec{u}) &= \mathcal{N}(  \nabla \vec{u} \mid 0, \mathbf{B}^{-1}),
  \label{eq:thikonov}
  \end{align}
%
  imposing that the distortion and its gradient have zero
  mean and variance governed by the matrices $\mathbf{A}$ and $\mathbf{B}$.
Finally, the maximum a posteriori problem is adapted into a variational one where we look for
  the minimum of an energy functional, by applying $E(\vec{u}) = -\log P(U)$:

  \begin{align}
  E(\vec{u}) &= -\log \underset{l}{\prod}
  \underset{\vec{r} \in \Omega_l}{\prod}
  \mathcal{N} \left( \vec{f}' \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_l \right)\,p_0( \vec{u})\,p_1( \vec{u}) \notag\\
  &= \const + \underset{l}{\sum} \int_{\Omega_l}
  \mdist{f'}{l} \,d\vec{r} \, + \notag\\
  &+ \int_{\Omega} \frac12 \left[ \vec{u}^T \mathbf{A} \vec{u} + (\nabla \vec{u})^T \mathbf{B} (\nabla \vec{u}) \right] \,d\vec{r}.
  \label{eq:energy}
  \end{align}
%
This expression is dual of the Mumford-Shah functional corresponding
  to a framework of \acrlong*{acwe} \citep{chan_active_2001}
  with the corresponding anisotropic regularization term of \cite{nagel_investigation_1986}.

@@

  \begin{align}
  \frac{\partial E(\vec{u})}{\partial \vec{u}_k} &=
  \frac{ \partial }{\partial \vec{u}_k} \Big\{
  \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} \, + \notag\\
  &+ \int_{\Omega} \frac12 [ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2}
  + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} ] \,d\vec{r}
  \Big\},
  \label{eq:gradient_descent}
  \end{align}
%
  where $\vec{u}^{\circ2} = \vec{u}^T \cdot \vec{u}$.
Then, the data and regularization terms are split and discretized to compute their
  derivatives.
The computation of the explicit shape gradients at each $\vec{v}_i'$ is described in \autoref{fig:method}.
Then, \eqref{eq:gradient_descent} can be reformulated as (see \suppl{equations SM6-SM12}):

@@

\paragraph*{Implementation details, settings and convergence}
\label{sec:conv_report}
The \emph{regseg} tool includes a multiresolution strategy on the free-form deformation field.
Registration pyramids are created by setting the spacing between control points of the B-Spline basis
  functions for each level of the multiresolution.
This implementation detail, along with other features (as the sparse matrix approach
  to fast interpolation) and the main parameters 
  (such as $\delta$, $\boldsymbol{\alpha}$, $\boldsymbol{\beta}$, the B-Spline grid resolutions,
  target image smoothing, etc.) are discussed in the \suppl{section S1}.
The actual choices of parameter settings are publicly distributed with the source code of the experiments.
These final settings were manually obtained based on the feedback of post-registration convergence
  reports (as the one found in the \suppl{section S1.3}).
We release \emph{regseg} along with the tool for generating the convergence report.

\subsection{Evaluation protocol}\label{sec:evaluation_protocol}
In order to assess the performance of \emph{regseg}, we defined a general
  evaluation protocol:
1) Extraction of the set of undistorted surfaces $\gammaset_R$;
2) computation of a ground-truth field of displacements $U_{true}$ that is applied to
  generate warped images ($M$) to segment;
3) execution of \emph{regseg} with $\gammaset_R$ and the warped data as inputs; and
4) visual assessment and computation of error metrics.
How this protocol is adapted to the simulated phantoms and real data is covered in the
  following sections.

\subsection{Simulated phantoms}\label{sec:digital_phantoms}
The workflow to simulate the digital phantoms and assess
  \emph{regseg} on them is presented in \autoref{fig:evphantoms}.
A set of four binary objects (called ``box'', ``ball'', ``L'',
  and ``gyrus'') are generated combining binarization of
  analytical shapes and mathematical morphology.
The reference surfaces $\gammaset_R$ are extracted from the binary shapes
  using \emph{FreeSurfer} tools \citep{fischl_freesurfer_2012}.
The ground-truth distortion is generated with a chain of two displacement 
  fields supported by grids of B-Spline basis functions.
The coefficients of the basis functions are generated randomly for
  both levels in their three dimensions.
The three components of the displacements $\vec{u} = (u_d)$ 
  are bounded above by 40\% of the separation between control points
  at each level, to obtain diffeomorphic transforms
  after concatenation \citep{rueckert_diffeomorphic_2006}.
The first deformation field is intended to generate large warpings,
  with control points separated by 50.50mm along the three dimensions
  ($u_d\leq$ 20.20mm).
With the second warping, we look for a resulting field with smoothness
  close to the found in a typical distortion field of \gls*{dmri} data
  \citep{irfanoglu_susceptibility_2011}.
Therefore, the control points are separated 25.25mm ($u_d\leq$ 10.10mm).
Once the ground-truth deformation is generated, the original surfaces
  are warped interpolating the displacements field at each vertex.
The warped surfaces $\gammaset_\text{true}$ are binarized to generate tissue fractions
  in low (\isores[mm]{2.0}) and high (\isores[mm]{1.0}) resolution.
Using a \gls*{mr} simulator \citep{caruyer_phantomas_2014}, we synthesize
  \gls*{t1} (TE/TR= 10/1500ms) and \gls*{t2} images (TE/TR= 90/5000ms)
  corresponding to each phantom type, with two resolutions each
  (1.0mm and 2.0mm isotropic).
The field of view at both resolutions is \isores[mm]{100}.
Then, \emph{regseg} is applied to map $\gammaset_R$ into the warped phantoms to
  obtain the registered surfaces ($\hat{\gammaset}_{test}$).
% Simulations were corrupted with rician noise for a \gls*{snr} of 300.0.
To quantify the misregistration error, we compute the Hausdorff distance between
 $\hat{\gammaset}_{test}$ and $\gammaset_\text{true}$, using \citep{commandeur_vtk_2011}.

\begin{figure*}
\includegraphics[width=\linewidth]{figure02}
\caption{The evaluation of \emph{regseg} on phantom data is performed with the presented instrumental workflow:
  1) The reference surfaces $\gammaset_R$ are triangularized meshes extracted from the four binary shapes (``box'', ``ball'', ``L'', ``gyrus'');
  2) A ground-truth displacement field is generated as described in \autoref{sec:digital_phantoms}, and applied to warp
      $\gammaset_R$, obtaining $\gammaset_\text{true}$;
  3) Once warped, $\gammaset_\text{true}$ are projected to the corresponding discrete 3D volume, and downsampled creating partial volume effects to two resolutions
     of \isores[mm]{2.0} and \isores[mm]{1.0}, producing sets of tissue fractions maps;
  4) The tissue fractions are fed to an \acrfull*{mr} simulator, generating \acrfull*{t1} and \acrfull*{t2} -like images at the
     two available resolutions;
  5) The \emph{regseg} is run, using the warped test images as moving multispectral-image and $\gammaset_R$ as shape-priors;
  6) Assessment of the agreement between the surfaces fitted with \emph{regseg} ($\hat{\gammaset}_{test}$) and $\gammaset_\text{true}$
     with automatically generated visual reports and computing the Hausdorff distance between corresponding surfaces.}\label{fig:evphantoms}
\end{figure*}

@@

\begin{figure*}
\includegraphics[width=\linewidth]{figure03}
\caption{Experimental workflow real data from the \acrfull*{hcp}.
  1) $\gammaset_R$ are extracted from the anatomical reference (\gls*{t1} image).
  2) To operate as ground truth, we generate a plausible-synthetic distortion $U_{true}$
    from the fieldmap using \eqref{eq:fieldmap}.
  3) The \gls*{dmri} data are warped using $U_{true}$ to reproduce the effects of real
    susceptibility-derived distortions.
  Target diffusion scalars (\gls*{fa} and \gls*{adc}) are computed on the distorted data and
    stacked to feed the multivariate input required by \emph{regseg}.
  4) the method is run, obtaining a $U_{test} = \hat{U}_{true}$, the estimation of
    the ground-truth deformation.
  5) Results are visually and quantitatively evaluated.}\label{fig:evworkflows}
\end{figure*}

\paragraph*{Data}
For the evaluation of \emph{regseg} on real \gls*{dmri} data of human brains,
  we collected 16 subjects from the \gls*{hcp} database.
The original acquisitions are released within ``unprocessed'' packages, and
  the ``minimally preprocessed'' are corrected for artifacts, brain-extracted
  and spatially normalized, along with some results of the standard processing
  pipeline of \emph{FreeSurfer}.
We refer the reader to \citep{essen_human_2012} for exact details about acquisition
  parameters, and \citep{glasser_minimal_2013} for the preprocessing issues.
The datasets comprehend a large set of images, containing \gls*{t1}, \gls*{t2} and
  multi-shell \gls*{dmri} images.

@@

  \begin{equation}
  sWI = \frac{1}{\sum_i a_i} \sum\limits_i^{N_V} a_i\,\|
  \vec{v}_i - \hat{\vec{v}}_i \|
  \label{eq:swindex}
  \end{equation}
%
  where $\vec{v}_i \subset \gammaset_\text{true}$ are the locations of the total $N_V$ vertices,
  and $\hat{\vec{v}}_i \subset \hat{\gammaset}_{test}$ are the locations
  recovered corresponding to $\vec{v}_i$.
In practice, we report the \gls*{swindex} only for three surfaces of interest crucial in whole-brain
  tractography.
These three surfaces delineate the following regions in the model: $\Omega_{VdGM}$, $\Omega_{WM}$ and $\Omega_{GM}$.

>>>>

 	\begin{align}
  P( M \mid U) &= \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod}
  \mathcal{N} ( \vec{f} \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} ) = % \notag\\ &=
  \underset{l}{\prod} \underset{\vec{r} \in \Omega_l}{\prod} \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{l}\right|}}\,{e^{\left(-\frac{1}{2}
  \mdist{f'}{l} \right)}},
  \label{eq:pdf}
  \end{align}
%
  using $\mdist{f}{l}$ to denote the squared \emph{Mahalanobis distance} of $\vec{f}$ with respect
  to the descriptors of region $l$ as
  $\mdist{f}{l} = (\vec{f} - \boldsymbol{\mu}_l)^T \, {\boldsymbol{\Sigma}_l}^{-1} \, (\vec{f} - \boldsymbol{\mu}_l)$.
$C$ is the number of channels comprised in the image $M$.
In fact, the projection of $\gammaset_R$ onto $M$ is an implicit segmentation model, for which
  the covariance matrix $\boldsymbol{\Sigma}_l$ of each region is minimized.

@@

  \begin{align}
  P(U) = \underset{\vec{r}}{\prod}\, p(\vec{u}) &=
  \underset{\vec{r}}{\prod}\, p_0(\vec{u}) \, p_1(\vec{u}), \text{ with} \\
  p_0(\vec{u}) &= \mathcal{N}( \vec{u} \mid 0, \mathbf{A}^{-1}), \notag\\
  p_1(\vec{u}) &= \mathcal{N}(  \nabla \vec{u} \mid 0, \mathbf{B}^{-1}),
  \label{eq:thikonov}
  \end{align}
%
  imposing that the distortion and its gradient have zero
  mean and variance governed by the matrices $\mathbf{A}$ and $\mathbf{B}$.
Finally, the maximum a posteriori problem is adapted into a variational one where we look for
  the minimum of an energy functional, by applying $E(\vec{u}) = -\log P(U)$:

  \begin{align}
  E(\vec{u}) &= -\log \underset{l}{\prod}
  \underset{\vec{r} \in \Omega_l}{\prod}
  \mathcal{N} \left( \vec{f}' \mid \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_l \right)\,p_0( \vec{u})\,p_1( \vec{u}) = \notag\\ &=
  \const + \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} \, + % \notag\\ &+
  \int_{\Omega} \frac12 \left[ \vec{u}^T \mathbf{A} \vec{u} + (\nabla \vec{u})^T \mathbf{B} (\nabla \vec{u}) \right] \,d\vec{r}.
  \label{eq:energy}
  \end{align}
%
This expression is dual of the Mumford-Shah functional corresponding
  to a framework of \acrlong*{acwe} \citep{chan_active_2001}
  with the corresponding anisotropic regularization term of \cite{nagel_investigation_1986}.

@@

  \begin{align}
  \frac{\partial E(\vec{u})}{\partial \vec{u}_k} &=
  \frac{ \partial }{\partial \vec{u}_k} \Big\{
  \underset{l}{\sum} \int_{\Omega_l} \mdist{f'}{l} \,d\vec{r} \, + % \notag\\ &+
  \int_{\Omega} \frac12 [ \boldsymbol{\alpha} \cdot \vec{u}^{\circ2}
  + \boldsymbol{\beta} \cdot (\nabla \vec{u})^{\circ2} ] \,d\vec{r}
  \Big\},
  \label{eq:gradient_descent}
  \end{align}
%
  where $\vec{u}^{\circ2} = \vec{u}^T \cdot \vec{u}$.
Then, the data and regularization terms are split and discretized to compute their
  derivatives.
The computation of the explicit shape gradients at each $\vec{v}_i'$ is described in \autoref{fig:method}.
Then, \eqref{eq:gradient_descent} can be reformulated as (see \suppl{equations SM6-SM12}):

@@

\paragraph*{Implementation details, settings and convergence}
\label{sec:conv_report}
The \regseg{} tool includes a multiresolution strategy on the free-form deformation field.
Registration pyramids are created by setting the spacing between control points of the B-Spline basis
  functions for each level of the multiresolution.
This implementation detail, along with other features (as the sparse matrix approach
  to fast interpolation) and the main parameters 
  (such as $\delta$, $\boldsymbol{\alpha}$, $\boldsymbol{\beta}$, the B-Spline grid resolutions,
  target image smoothing, etc.) are discussed in the \suppl{section S1}.
The actual choices of parameter settings are publicly distributed with the source code of the experiments.
These final settings were manually obtained based on the feedback of post-registration convergence
  reports (as the one found in the \suppl{section S1.3}).
We release \regseg{} along with the tool for generating the convergence report.

\subsection{Evaluation protocol}\label{sec:evaluation_protocol}
In order to assess the performance of \regseg{}, we defined a general
  evaluation protocol:
1) Extraction of the set of undistorted surfaces $\gammaset_R$;
2) computation of a ground-truth field of displacements $U_{true}$ that is applied to
  generate warped images ($M$) to segment;
3) execution of \regseg{} with $\gammaset_R$ and the warped data as inputs; and
4) visual assessment and computation of error metrics.
How this protocol is adapted to the simulated phantoms and real data is covered in the
  following sections.

\subsection{Simulated phantoms}\label{sec:digital_phantoms}
The workflow to simulate the digital phantoms and assess
  \regseg{} on them is presented in \autoref{fig:evphantoms}.
A set of four binary objects (called ``box'', ``ball'', ``L'',
  and ``gyrus'') are generated combining binarization of
  analytical shapes and mathematical morphology.
The reference surfaces $\gammaset_R$ are extracted from the binary shapes
  using \emph{FreeSurfer} tools \citep{fischl_freesurfer_2012}.
The ground-truth distortion is generated with a chain of two displacement 
  fields supported by grids of B-Spline basis functions.
The coefficients of the basis functions are generated randomly for
  both levels in their three dimensions.
The three components of the displacements $\vec{u} = (u_d)$ 
  are bounded above by 40\% of the separation between control points
  at each level, to obtain diffeomorphic transforms
  after concatenation \citep{rueckert_diffeomorphic_2006}.
The first deformation field is intended to generate large warpings,
  with control points separated by 50.50mm along the three dimensions
  ($u_d\leq$ 20.20mm).
With the second warping, we look for a resulting field with smoothness
  close to the found in a typical distortion field of \gls*{dmri} data
  \citep{irfanoglu_susceptibility_2011}.
Therefore, the control points are separated 25.25mm ($u_d\leq$ 10.10mm).
Once the ground-truth deformation is generated, the original surfaces
  are warped interpolating the displacements field at each vertex.
The warped surfaces $\gammaset_\text{true}$ are binarized to generate tissue fractions
  in low (\isores[mm]{2.0}) and high (\isores[mm]{1.0}) resolution.
Using a \gls*{mr} simulator \citep{caruyer_phantomas_2014}, we synthesize
  \gls*{t1} (TE/TR= 10/1500ms) and \gls*{t2} images (TE/TR= 90/5000ms)
  corresponding to each phantom type, with two resolutions each
  (1.0mm and 2.0mm isotropic).
The field of view at both resolutions is \isores[mm]{100}.
Then, \regseg{} is applied to map $\gammaset_R$ into the warped phantoms to
  obtain the registered surfaces ($\hat{\gammaset}_{test}$).
% Simulations were corrupted with rician noise for a \gls*{snr} of 300.0.
To quantify the misregistration error, we compute the Hausdorff distance between
 $\hat{\gammaset}_{test}$ and $\gammaset_\text{true}$, using \citep{commandeur_vtk_2011}.

\begin{figure*}
\includegraphics[width=\linewidth]{figure02}
\caption{The evaluation of \regseg{} on phantom data is performed with the presented instrumental workflow:
  1) The reference surfaces $\gammaset_R$ are triangularized meshes extracted from the four binary shapes (``box'', ``ball'', ``L'', ``gyrus'');
  2) A ground-truth displacement field is generated as described in \autoref{sec:digital_phantoms}, and applied to warp
      $\gammaset_R$, obtaining $\gammaset_\text{true}$;
  3) Once warped, $\gammaset_\text{true}$ are projected to the corresponding discrete 3D volume, and downsampled creating partial volume effects to two resolutions
     of \isores[mm]{2.0} and \isores[mm]{1.0}, producing sets of tissue fractions maps;
  4) The tissue fractions are fed to an \acrfull*{mr} simulator, generating \acrfull*{t1} and \acrfull*{t2} -like images at the
     two available resolutions;
  5) The \regseg{} is run, using the warped test images as moving multispectral-image and $\gammaset_R$ as shape-priors;
  6) Assessment of the agreement between the surfaces fitted with \regseg{} ($\hat{\gammaset}_{test}$) and $\gammaset_\text{true}$
     with automatically generated visual reports and computing the Hausdorff distance between corresponding surfaces.}\label{fig:evphantoms}
\end{figure*}

@@

\begin{figure*}
\includegraphics[width=\linewidth]{figure03}
\caption{Experimental workflow real data from the \acrfull*{hcp}.
  1) $\gammaset_R$ are extracted from the anatomical reference (\gls*{t1} image).
  2) To operate as ground truth, we generate a plausible-synthetic distortion $U_{true}$
    from the fieldmap using \eqref{eq:fieldmap}.
  3) The \gls*{dmri} data are warped using $U_{true}$ to reproduce the effects of real
    susceptibility-derived distortions.
  Target diffusion scalars (\gls*{fa} and \gls*{adc}) are computed on the distorted data and
    stacked to feed the multivariate input required by \regseg{}.
  4) the method is run, obtaining a $U_{test} = \hat{U}_{true}$, the estimation of
    the ground-truth deformation.
  5) Results are visually and quantitatively evaluated.}\label{fig:evworkflows}
\end{figure*}

\paragraph*{Data}
For the evaluation of \regseg{} on real \gls*{dmri} data of human brains,
  we collected 16 subjects from the \gls*{hcp} database.
The original acquisitions are released within ``unprocessed'' packages, and
  the ``minimally preprocessed'' are corrected for artifacts, brain-extracted
  and spatially normalized, along with some results of the standard processing
  pipeline of \emph{FreeSurfer}.
We refer the reader to \citep{essen_human_2012} for exact details about acquisition
  parameters, and \citep{glasser_minimal_2013} for the preprocessing issues.
The datasets comprehend a large set of images, containing \gls*{t1}, \gls*{t2} and
  multi-shell \gls*{dmri} images.

@@

  \begin{equation}
  sWI = \frac{1}{\sum_i a_i} \sum\limits_i^{N_V} a_i\,\|
  \vec{v}_i - \hat{\vec{v}}_i \|,
  \label{eq:swindex}
  \end{equation}
%
  where $\vec{v}_i \subset \gammaset_\text{true}$ are the locations of the total $N_V$ vertices,
  and $\hat{\vec{v}}_i \subset \hat{\gammaset}_{test}$ are the locations
  recovered corresponding to $\vec{v}_i$.
In practice, we report the \gls*{swindex} only for three surfaces of interest crucial in whole-brain
  tractography.
These three surfaces delineate the following regions in the model: $\Omega_{VdGM}$, $\Omega_{WM}$ and $\Omega_{GM}$.
